Pottery decor as networks on the Middle Niger

### 2024-03-08 analysis/paper/paper.Rmd
script in proper order, workin iNext, communities based on sparsified networks and better community comparison

```{r backbone-extraction, fig.cap="Extraction of binary backbones at decreasing significance levels: a) 0.05, b) 0.15, c) 0.18"}
# extract binary backbones at decreasing levels of significance
decor_05 <- extract_backbone(decor_flat, 0.05)
decor_15 <- extract_backbone(decor_flat, 0.15)
decor_18 <- extract_backbone(decor_flat, 0.20) #20 ideale anzahl?

# plot
p1 <- plot.bb(decor_05)
p2 <- plot.bb(decor_15)
p3 <- plot.bb(decor_18) + geom_node_text(aes(label = label), repel = TRUE)

@@

decor_18 <- decor_18 |>
  #isolated nodes are removed
  activate(nodes) |>
  filter(!node_is_isolated())

@@

```{r backbone-eigenvalue, fig.cap="Binary backbone at alpha = 0.15 showing eigenvector centralities"}
ggraph(decor_18, layout = "nicely") +
  geom_node_point(aes(size = centrality_eigen())) +
  geom_edge_fan(color = "grey", alpha = 0.8) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "eigenvector centrality")
```

@@

```{r backbone-betweenness, fig.cap="Binary backbone at alpha = 0.15 showing betweenness centralities"}
ggraph(decor_18, layout = "nicely") +
  geom_node_point(aes(size = centrality_betweenness())) +
  geom_edge_fan(color = "grey", alpha = 0.8) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "betweenness centrality")
```

@@

```{r community detection, hide = 'results', echo=FALSE}
decor_graph1 <- decor_20 |>
  activate(nodes) |>
  mutate(louvain = as.character(group_louvain()))

decor_graph1 <- decor_graph |>
  activate(nodes) |>
  mutate(louvain = as.data.frame(decor_graph1, "vertices")$louvain)

# calculate modularity of louvain clustering
modularity <- decor_graph1 |>
  activate(nodes) |>
  mutate(modularity = graph_modularity(group = as.factor(louvain))) |>
  pull(modularity) |>
  head(1)

decor_graph1

decor_graph |>
  activate(edges) |> 
  filter(str_detect(name, "roulette") | name %in% c("paint", "straight_mat_impression"))  |> 
  as.data.frame("edges") |> 
  group_by(name) |> 
  summarise()

decor_graph |>
  activate(edges) |> 
  filter(str_detect(name, "strip_roulette")| str_detect(name, "cord_roulette"))  |> 
  as.data.frame("edges") |> 
  group_by(name) |> 
  summarise()

decor_flat
```

@@

```{r plot-clusters, fig.cap="Louvain communities within the decor network", echo = FALSE}
# set plot colours for communities
communitycolor <-
  setNames(
    c("tomato", "cornflowerblue", "darkgreen", "yellow"),
    unique(V(decor_graph1)$louvain)
  )

@@

ggraph(decor_graph1, x = lon, y = lat) +
  geom_sf(data = mali_adm0, fill = NA) +
  geom_sf(data = rivers, colour = "black") +
  edges_community +
  ggforce::geom_mark_hull(aes(x = lon, y = lat,
                              fill = louvain, colour = louvain),
                          expand = unit(1.5, "mm"),
                          radius = unit(1.5, "mm")) +
  scale_fill_manual(values = communitycolor) +
  scale_colour_manual(values = communitycolor) +
  nodes_community +
  theme_community +
  coord_sf(xlim = c(-7, 2), ylim = c(13, 19), expand = FALSE) +
  theme_graph()

@@

```{r cluster1, fig.cap="Internal connections within community 1"}
decor_graph1 %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()

decor_18 %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()

decor_flat %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster2, fig.cap="Internal connections within community 2"}
decor_graph1 %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 2) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "tomato") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph1 %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "darkgreen") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```
```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph1 %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 4) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "yellow") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster-comparison, fig.cap="Relative edge weights and centralization by cluster"}
# adding community/crossing to the edges
decor_graph1 <- decor_graph1 |>
 # flatten(weighted = TRUE) |> 
  add_community_edges()

decor_graph1

# mean edge weights
mean_edge_weight <- decor_graph1 |>
  get.data.frame("edges") |>
#  filter(community != "crossing") |>
  group_by(from,to, community) |> 
  summarise(weight = sum(weight),
  div = n()) #|> 
 # group_by(community) |>
 # summarise(median_weight = median(weight),
  #          mean_weight = mean(weight))

# Centralization of the cluster
centralization <- vector("double", length(4))
for (i in 1:4) {
  centralization[[i]] <- decor_graph1 |>
    # filter community
    activate(nodes) |>
    filter(louvain == i) |>
    # degree centralization
    centr_degree(mode = "all", normalized = TRUE) |>
    as_tibble() |>
    pull(centralization) |>
    head(1)
}

# Graph of mean edge weight of the different communities
g1 <- mean_edge_weight |>
  pivot_longer(!community, names_to = "key",
               values_to = "mean_edge_weight") |>
  ggplot(aes(x = community, y = mean_edge_weight)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_grey() +
  theme_bw() +
  ylab("mean edge weight")

# Graph of centralization of the different communities
g2 <- tibble(centralization) |>
  mutate(community = 1:4) |>
  pivot_longer(!community, names_to = "key", values_to = "centralisation") |>
  ggplot(aes(x = community, y = centralisation)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_grey() +
  theme_bw()

# plot:
g1 + g2

ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = weight, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "degree centrality",
       x = "community")

ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = div, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "degree centrality",
       x = "community")

ggplot(centrali) +
  boxplot(aes(x = community, y = div, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "degree centrality",
       x = "community")
```

@@

# Calculate centralities for each cluster
centr <- pull_centralities_clust(decor_graph1)

@@

#show crossings with a weight over 3
decor_graph1 |>
  filter(
    community == "crossing" &
    weight > 3 
  # & !name %in% c("channel", "twisted_cord_roulette")
  ) |>
  activate(nodes) |>
  filter(!node_is_isolated()) |>
  #plot
  ggraph(layout = "stress") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight),
                      color = louvain), show.legend = FALSE) +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  scale_color_manual(values = communitycolor) +
  theme_graph() +
  theme_community
```

@@

library(backbone)
# extract backbone
early_flat_bb <- 
  backbone::disparity(as.igraph(early_flat), alpha = 0.22, narrative = TRUE) |>
  tidygraph::as_tbl_graph()

>>>>

```{r backbone-extraction, fig.cap="Extraction of binary backbones at decreasing significance levels: a) 0.05, b) 0.15, c) 0.18"}
# extract binary backbones at decreasing levels of significance
decor_05 <- extract_backbone(decor_flat, 0.05)
decor_15 <- extract_backbone(decor_flat, 0.15)
decor_20 <- extract_backbone(decor_flat, 0.20) #20 ideale anzahl?

# plot
p1 <- plot.bb(decor_05)
p2 <- plot.bb(decor_15)
p3 <- plot.bb(decor_20) + geom_node_text(aes(label = label), repel = TRUE)

@@

decor_20 <- decor_20 |>
  #isolated nodes are removed
  activate(nodes) |>
  filter(!node_is_isolated())

@@

```{r backbone-eigenvalue, fig.cap="Binary backbone at alpha = 0.15 showing eigenvector centralities"}
ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_eigen())) +
  geom_edge_fan(color = "grey", alpha = 0.8) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "eigenvector centrality")
```

@@

```{r backbone-betweenness, fig.cap="Binary backbone at alpha = 0.15 showing betweenness centralities"}
ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_betweenness())) +
  geom_edge_fan(color = "grey", alpha = 0.8) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "betweenness centrality")
```

@@

```{r community detection, hide = 'results', echo=FALSE}
#sparsifying 
decor_graph_c <- decor_20 |>
  activate(nodes) |>
  mutate(louvain = as.character(group_louvain()))

decor_graph <- decor_graph |>
  activate(nodes) |>
  mutate(louvain = as.data.frame(decor_graph_c, "vertices")$louvain)

# calculate modularity of louvain clustering
modularity <- decor_graph |>
  activate(nodes) |>
  mutate(modularity = graph_modularity(group = as.factor(louvain))) |>
  pull(modularity) |>
  head(1)
```

@@

```{r plot-clusters, fig.cap="Louvain communities within the decor network", echo = FALSE}
# set plot colours for communities
communitycolor <-
  setNames(
    c("cornflowerblue","tomato", "darkgreen", "yellow"),
    unique(V(decor_graph)$louvain)
  )

@@

ggraph(decor_graph, x = lon, y = lat) +
  geom_sf(data = mali_adm0, fill = NA) +
  geom_sf(data = rivers, colour = "black") +
  edges_community +
  ggforce::geom_mark_hull(aes(x = lon, y = lat,
                              fill = louvain, colour = louvain),
                          expand = unit(1.5, "mm"),
                          radius = unit(1.5, "mm")) +
  scale_fill_manual(values = communitycolor) +
  scale_colour_manual(values = communitycolor) +
  nodes_community +
  theme_community +
  coord_sf(xlim = c(-7, 2), ylim = c(13, 19), expand = FALSE) +
  theme_graph()

@@

```{r cluster1, fig.cap="Internal connections within community 1"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster2, fig.cap="Internal connections within community 2"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 2) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "tomato") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "darkgreen") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```
```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 4) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "yellow") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster-comparison, fig.cap="Relative edge weights, edge diversity and centralization by cluster"}
# adding community/crossing to the edges
decor_graph <- decor_graph |>
 # flatten(weighted = TRUE) |> 
  add_community_edges()

# mean edge weights
mean_edge_weight <- decor_graph |>
  get.data.frame("edges") |>
#  filter(community != "crossing") |>
  group_by(from,to, community) |> 
  summarise(weight = sum(weight),
  div = n()) #|> 
 # group_by(community) |>
 # summarise(median_weight = median(weight),
  #          mean_weight = mean(weight))

# Centralization of the cluster
centralization <- vector("double", length(4))
for (i in 1:4) {
  centralization[[i]] <- decor_graph |>
    # filter community
    activate(nodes) |>
    filter(louvain == i) |>
    # degree centralization
    centr_degree(mode = "all", normalized = TRUE) |>
    as_tibble() |>
    pull(centralization) |>
    head(1)
}

# Graph of mean edge weight of the different communities
g1 <- ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = weight, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "degree centrality",
       x = "community")

# Graph of centralization of the different communities
g2 <- ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = div, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "degree centrality",
       x = "community")

#Graph of centralization of the communities
g3 <- tibble(centralization) |>
  mutate(community = 1:4) |>
  pivot_longer(!community, names_to = "key", values_to = "centralisation") |>
  ggplot(aes(x = community, y = centralisation)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_grey() +
  theme_bw()

# plot:
g1 + g2 +g3
```

@@

# Calculate centralities for each cluster
centr <- pull_centralities_clust(decor_graph)

@@

#show crossings with a weight over 3
decor_graph |>
  filter(
    community == "crossing" &
    weight > 3 
  # & !name %in% c("channel", "twisted_cord_roulette")
  ) |>
  activate(nodes) |>
  filter(!node_is_isolated()) |>
  #plot
  ggraph(layout = "stress") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight),
                      color = louvain), show.legend = FALSE) +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  scale_color_manual(values = communitycolor) +
  theme_graph() +
  theme_community
```

@@

library(backbone)
# extract backbone
early_flat_bb <- 
  backbone::disparity(as.igraph(early_flat), alpha = 0.21, narrative = TRUE) |>
  tidygraph::as_tbl_graph()
